---
title: "Predictors of Testing Positive for Diabetes"
author: Daniel Molitor
format:
  html:
    code-tools:
      source: https://github.com/dmolitor/nextflow-example
    theme: darkly
    embed-resources: false
params:
  data_path: ""
  oos_error_path: ""
  model_path: ""
---

```{r}
#| echo: false
#| label: Script Setup
#| warning: false
diabetes <- read.csv(params$data_path) |>
      transform(diabetes = factor(diabetes, levels = c("neg", "pos")))
    oos_error <- round(100*as.numeric(readLines(params$oos_error_path)), 2)
    best_tree <- readRDS(params$model_path)
```

This analysis uses the [Pima Indians Diabetes dataset](http://math.furman.edu/~dcs/courses/math47/R/library/mlbench/html/PimaIndiansDiabetes.html)
to determine which factors are important predictors of testing positive for
diabetes.

## Data Structure

Let's get a quick understanding of the variables included in our dataset
and how they relate to each other.
```{r}
#| label: Data Structure
#| layout: [[100]]
#| warning: false
str(diabetes)

plot(diabetes, col = "lightblue")
```

## Optimal Decision Tree Model

The optimal Decision Tree (hold-out set accuracy of 79.22%) model
has the following hyper-parameters:
```{r}
#| label: Optimal Tree
#| warning: false
jsonlite::toJSON(best_tree$control, pretty = TRUE, auto_unbox = TRUE)
```

We can also plot the tree structure using `rpart.plot`.
```{r}
#| label: Tree Structure
#| layout: [[100]]
#| warning: false
rpart.plot::rpart.plot(best_tree, roundint = FALSE)
```

## Variable Importance

Finally, we can determine the most important predictors of a positive
diabetes diagnosis, using the variable importance feature of our
Decision Tree.
```{r}
#| label: Variable Importance
#| layout: [[100]]
#| warning: false
var_importance <- transform(
  data.frame(
    "Feature" = names(best_tree$variable.importance),
    "Importance" = best_tree$variable.importance
  ),
  Importance = Importance/sum(Importance)
)

ggplot2::ggplot(
  var_importance,
  ggplot2::aes(x = forcats::fct_reorder(Feature, Importance), y = Importance)
) +
  ggplot2::geom_bar(stat = "identity") +
  ggplot2::labs(x = "Feature") +
  ggplot2::theme_light() +
  ggplot2::theme(
    axis.title = ggplot2::element_text(face = "bold"),
    panel.border = ggplot2::element_blank()
  )
```

## Conclusion

Unsurprisingly, glucose levels are by far the most important indicator of
diabetes.